---
title: "Bike Trip"
author: "Me"
date: "2024-07-09"
output: html_document
---
Cyclistic is a bike-sharing operation located in Chicago. Cyclistic would like to convert casual riders into annual members. Our goal for this analysis is to answer the question: How do annual member and casual riders use Cyclistic bikes differently.  

Our data consists of rider data from a bike sharing service for the first quarters of the years 2019 and 2020. The data has been made available by Motivate International Inc. under this [liscense](https://divvybikes.com/data-license-agreement).  

## Exploratory Analysis in Sheets
  
  * Created trip length and day of the week columns
  * Cleaned up column names to make naming consistent between the two data     sets
  * Data initially consisted of 14 columns for the 2019 data and 15 columns     for the 2020 data. After removing superfluous columns, both data sets
    have 10 columns.
  * Changed Customer under user_type in 2019 data to casual and Subscriber     to member.
  
## Beginning R Analysis

```{r Packages}
library(tidyverse)
library(skimr)
library(janitor)
```

```{r Importing Data}
bike_data_2019 <- read.csv('bike_data_2019_Q1.csv')
bike_data_2020 <- read.csv('bike_data_2020_Q1.csv')
```

Let's start by combining our two data sets into a single data frame and converting the trip_length data to numeric data type.

```{r Aggregating Data}
df_bike_data <- merge(x = bike_data_2019, y = bike_data_2020, all = TRUE)

df_bike_data[["trip_length"]] <- as.duration(hms(df_bike_data[["trip_length"]]))
```

Our data contains a number of rides of negligible length and other rides of several weeks. We will remove these outliers to get a more accurate picture of the average user.

```{r Removing Outliers}
df_trimmed <- df_bike_data %>%
  filter(trip_length >60 & trip_length < (12*3600))
```

To start our analysis let's calculate the average trip length for members and casual riders measured in minutes.

```{r}
df_trimmed %>%
  group_by(user_type) %>%
  summarise(mean = mean(as.numeric(trip_length)/60))
```

As we can see, casual riders are on average taking much longer rides than members. Let's continue by finding the mean for each day of the week.

```{r}
df_trimmed %>%
  group_by(day_of_week, user_type) %>%
  summarise(mean = mean(as.numeric(trip_length)/60))
```

Here our day runs from 1 (Sunday) through 7 (Saturday).The average trip length does not vary greatly by day of the week. Let's obtain the number of riders by day of the week.

```{r}
df_trimmed %>%
  group_by(day_of_week, user_type) %>%
  summarise(riders = n())
```

Here we can see that while casual riders are riding more on the weekends, members are riding much more during the week.


## Conclusions

Our analysis showed that members on average take much shorter rides than casual riders and do so during the work week. We conclude that members are using bikes as their day-to-day transportation. Casual riders could be persuaded to become members if they too can see the utility of using the service for day-to-day transport. 

The shorter ride length of members is also interesting. Casual riders could be persuaded to become members if they can be shown the utility of using the service for short trips, perhaps by showing how quick it is to rent and return a bike as a member.
